---
title: "midvikudagur4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(tidyverse)

# 1)
```{r}
as <- read.table("https://notendur.hi.is/ahj/pulsAll.csv", sep = ";", header = T)

```

# 2)
```{r}
glimpse(as)
```

# 3)
```{r}
as <- mutate(as, kyn = factor(kyn))
as <- mutate(as, kyn = fct_recode(kyn, kvk = "1", kk = "2"))
```

# 4)
```{r}
as <- as %>% mutate(hreyfing = ifelse(as$likamsraekt > 0, "ja", "nei"))

```


# 5)
```{r}
set.seed(37) 
urtak <-sample_n(as, 200)
```



# 6)
```{r}
binom.test(table(urtak$hreyfing))

```
mörk: 0.8126270 < p < 0.9119421


# 7)
```{r}
binom.test(table(urtak$hreyfing))

```

data:  table(urtak$hreyfing)
number of successes = 171, number of trials = 197, p-value < 2.2e-16
alternative hypothesis: true probability of success is not equal to 0.5
95 percent confidence interval:
  0.8126270 0.9119421
sample estimates:
probability of success 
0.8680203 

h0: probability of success is 0.5
h1: probability of success is not 0.5
prófstærðin: 0.8680203
p-gildi: 2.2e-16
þar sem útkoman er 0.86 og öryggisbil er 0.81 - 0.911 getum við með 95% öryggi
hafnað núlltilgátunni.

# 8)
```{r}
prop.test(table(urtak$drekkur,urtak$namskeid))
```
data:  table(urtak$drekkur, urtak$namskeid)
X-squared = 2.3543, df = 1, p-value = 0.1249
alternative hypothesis: two.sided
95 percent confidence interval:
  -0.01477881  0.32187415
sample estimates:
  prop 1    prop 2 
0.3353659 0.1818182 

hér er p-gildið hærra en alpha og því samþykkjum við núll-tilgátuna
